---
title: '<img src="https://drive.google.com/uc?id=118y5T73-iSUZoAMtoJGddxq9QzD_GDKX" style="height:40px;float:left;margin:10px" /><img src="https://drive.google.com/uc?id=1fHQUzF3ZjaoHj9KQ33-94dK_X1hcmjzW" style="height:50px;float:left;margin:10px" />'

pagetitle: 'AA Trigger Report'
output:
  html_document:
    css: country_template/docs/style.css
    includes:
      in_header: country_template/docs/header.html
    toc: no
    toc_float: yes
    toc_depth: '4'
    df_print: paged
  html_notebook:
    css: country_template/docs/style.css
    includes:
      in_header: country_template/docs/header.html
    toc: no
    toc_float: yes
    toc_depth: '4'
    df_print: paged
editor_options:
  chunk_output_type: console
---
<style type="text/css">
.main-container {
  max-width: 90%;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r README, echo = FALSE, message = FALSE, warning = FALSE}
### NOTE: Performance metrics and their CIs should have been generated separately before the trigger report can be produced.

### Create a trigger_report folder under DATA_DIR > /private/exploration/{country_pcode3/trigger_performance}
### Create a "plots" folder sister to the RMD file in the country's `analyses` folder
### Update the "hard-coded-values" code chunk below
### If there are fewer or more than 3 triggers: Adjust the number of triggers and/or Packages throughout the Rmd (review the snapshot table and activation table carefully.)
### Review, adjust or remove the country-specific code chunk
### Knit
```

```{r hard-coded-values, echo = FALSE, message = FALSE, warning = FALSE}
country_name <- "The Philippines"
country_pcode3 <- "phl"
shock <- "Typhoons"

# trigger A info
name_a <- "readiness" # what name is the trigger listed as in the perf_metrics_table.csv files?
lead_time_a<- '4-7 days before landfall' # eg: 'Jan, Feb, Mar'
data_source_a <- 'PAGASA, ECMWF, GEFS, Weather Nerds' # name of product + provider
hist_period_a <- "2006 - 2021" # period for which historical data is available
monitored_by_a <- 'OCHA Philippines' # owner of monitoring responsibilities

# trigger B info
name_b <- "activation"
lead_time_b <- '72 hours before landfall'
data_source_b <- 'ECMWF, 510 model'
hist_period_b <- "2006 - 2021"
monitored_by_b <- 'OCHA Philippines'

# trigger C info
name_c <- "framework"
lead_time_c <- "N/A"
data_source_c <- 'N/A'
hist_period_c <- "N/A"
monitored_by_c <- 'N/A'

# Table footnotes
trigger_table_footnote <- "The readiness and activation trigger can reach their threshold and activate independently from one another. The full framework activation is met if both the readiness and activation triggers are reachedt. The readiness trigger performance was assessed by comparing recommended activations with historical typhoon forecasts, and the activation trigger by running the 510 model on ECMWF hindcasts."
#
# # Data providers
data_providers <- "OCHA Philippines, 510 Global"
#
# # Analysis
# analysis <- "Historical data and analysis are available through IRI's Maproom [tool](http://iridl.ldeo.columbia.edu/fbfmaproom2/niger). The triggers were developed by IRI in close collaboration with DMN, WFP Niger, and the Centre for Humanitarian Data with valuable input from participating agencies and OCHA."
#
# Monitoring process
monitoring_process <- "Between September and January, OCHA Philippines
monitors typhoon forecasts from the Philippine Atmospheric, Geophysical
and Astronomical Services Administration (PAGASA), the European Centre
for Medium-Range Weather Forecasts (ECMWF), the Global Ensemble Forecast
System (GEFS), and Weather Nerds (Private Initiative). If the typhoon
is projected to impact areas in regions 5, 8 and 13 within 4-7 days, and exceed 178
km/h 1-minute sustained wind speed, then the readiness trigger is reached, with
either scenario 1 or 2 specified. 72 hours before landfall, the 510 damage
assessment model is run. If any of the damage threshold numbers exceed
those specified in the framework by the target probability
(ranging from 50% probability that
80,000 houses or more will be totally damaged, or 95% probability
that 5,000 will be totally damaged), then the activation trigger is reached.
Note that for a successful activation, both the readiness and activation
trigger need to be reached."
```

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE) # do not print code by default
knitr::opts_chunk$set(include = TRUE) # include chunk output by default
knitr::opts_chunk$set(message = FALSE) # do not print messages by default
knitr::opts_chunk$set(warning = FALSE) # do not print warnings by default

options(scipen = 999) # turn off scientific notation
options(encoding = "UTF-8") # set encoding to UTF-8 instead of ANSI
```

```{r libraries, include = FALSE, warning = FALSE}
packages <- c('tidyverse', 'ggthemes', 'kableExtra', 'knitr', 'flextable')

# install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())

if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# load libraries
lapply(packages, library, character.only = TRUE)

```


```{r paths}
data_folder_path <- paste0(Sys.getenv("AA_DATA_DIR"), "/public/exploration/", country_pcode3,"/trigger_performance/")
perf_metrics_95_filename <- paste0(data_folder_path, country_pcode3, "_perf_metrics_table_ci_0.95.csv")
perf_metrics_68_filename <- paste0(data_folder_path, country_pcode3, "_perf_metrics_table_ci_0.68.csv")
learning_table_filename <- paste0(data_folder_path, country_pcode3, "_learning_table.csv")
explanatory_ci_plot_path <- ("country_template/docs/explanatory_ci.png")
```

```{r functions}
source("country_template/docs/plotAtvCI.R")
source("country_template/docs/plotTradeoffCI.R")
```

```{r data-loading-and-formatting}

# read in performance metrics files
perf_metrics_95_data <- read.csv(perf_metrics_95_filename,
                              colClasses = c('character', 'character', 'numeric', 'character')
                              )

perf_metrics_68_data <- read.csv(perf_metrics_68_filename,
                              colClasses = c('character', 'character', 'numeric', 'character')
                              )

# format performance metrics values
perf_metrics_95_data$value <- round(100 * perf_metrics_95_data$value, 0)
perf_metrics_68_data$value <- round(100 * perf_metrics_68_data$value, 0)

# uniquely identify point labels
perf_metrics_95_data$upoint <- paste0(perf_metrics_95_data$point, "_95")
perf_metrics_68_data$upoint <- paste0(perf_metrics_68_data$point, "_68")

# combine datasets
perf_metrics_data <- rbind(perf_metrics_95_data, perf_metrics_68_data) %>%
                      select(-point) # remove non-unique point identifier


```


```{r atv-CIs}

# Create a CI visual for atv for each trigger, activation timepoint, and activation scenario beside None and Any

trigger_list = c("readiness", "activation", "framework")

for(trigger in trigger_list) {

  # produce segment_dimensions and plots for triggers
  output <- plotAtvCI(trigger_id = trigger)

  # save plot as png
  filename <- paste0(trigger, "_atv.png")
  png(filename = paste0("plots/", filename), width = 500, height = 400, units = "px")
  print(output$p)
  invisible(dev.off())
}

```

```{r tradeoff-CIs}

# Tradeoff CIs for individual triggers

for(trigger in c(trigger_list)) {
  for (metric in c('var', 'det')) {

  output <- plotTradeoffCI(trigger_id = trigger,
                   left_metric_name = metric)

  # save plot as png
  filename <- paste0(trigger, "_", metric, "_tradeoff_bar.png")
  png(filename = paste0("plots/", filename), width = 500, height = 400, units = "px")
  print(output$p)
  invisible(dev.off())
  }
}

```

```{r file-formatting-and-saving}

# list plots to be trimmed
plots_to_trim <- grep(list.files(path = "./plots"),
                      pattern = 'trimmed_',
                      invert = TRUE,
                      value = TRUE) # excludes 'trimmed_' graphs, which will be overwritten

# trim plots
for (filename in plots_to_trim) {

  # read in the image
  original_plot <- magick::image_read(paste0("plots/", filename))

  # trim the image
  trimmed_plot <- magick::image_trim(original_plot)

  # save trimmed plot
  magick::image_write(trimmed_plot, path = paste0("plots/trimmed_", filename), format = "png")

}

```

```{r key-messages-metrics}
any_atv_low_end_68 <- perf_metrics_data %>% filter(trigger == 'framework' & metric == 'atv') %>% filter(upoint == 'low_end_68') %>% select(value) %>% as.numeric()

any_atv_high_end_68 <- perf_metrics_data %>% filter(trigger == 'framework' & metric == 'atv') %>% filter(upoint == 'high_end_68') %>% select(value) %>% as.numeric()

any_det_low_end_68 <- perf_metrics_data %>% filter(trigger == 'framework' & metric == 'det') %>% filter(upoint == 'low_end_68') %>% select(value) %>% as.numeric()

any_det_high_end_68 <- perf_metrics_data %>% filter(trigger == 'framework' & metric == 'det') %>% filter(upoint == 'high_end_68') %>% select(value) %>% as.numeric()

any_far_low_end_68 <- perf_metrics_data %>% filter(trigger == 'framework' & metric == 'far') %>% filter(upoint == 'low_end_68') %>% select(value) %>% as.numeric()

any_far_high_end_68 <- perf_metrics_data %>% filter(trigger == 'framework' & metric == 'far') %>% filter(upoint == 'high_end_68') %>% select(value) %>% as.numeric()

any_mis_low_end_68 <- perf_metrics_data %>% filter(trigger == 'framework' & metric == 'mis') %>% filter(upoint == 'low_end_68') %>% select(value) %>% as.numeric()

any_mis_high_end_68 <- perf_metrics_data %>% filter(trigger == 'framework' & metric == 'mis') %>% filter(upoint == 'high_end_68') %>% select(value) %>% as.numeric()


mitigation_warranted <- ifelse(any_far_low_end_68 >= 40, "Mitigation is recommended to minimise the negative impact of a missed activation due to its moderate to high likelihood.", "")

```

# Anticipatory Action Trigger Mechanism Report
<p style="color:#007ce1;font-size:22px;text-align:center;"> `r paste0(shock, ' in ', country_name)`</p>
<p style="color:black;font-size:14px;text-align:right;"> Last updated: `r Sys.Date()`</p>

## Key Messages

* <span style="color: blue;">ACTIVATION</span> <p style="color:black;font-size:16px;text-align:left;"> There is a `r paste0(any_atv_low_end_68, ' - ', any_atv_high_end_68)`% chance that a full activation (both readiness and activation trigger) will occur in a given year.</p>
* <span style="color: blue;">MISSED ACTIVATIONS</span> <p style="color:black;font-size:16px;text-align:left;"> The risk of not activating in the presence of a severe shock stands at `r paste0(any_mis_low_end_68, ' - ', any_mis_high_end_68)`% of shocks.</p>
* <span style="color: blue;">FALSE ALARMS</span> <p style="color:black;font-size:16px;text-align:left;"> The risk of activating in the absence of a severe shock stands at `r paste0(any_far_low_end_68, ' - ', any_far_high_end_68)`% of activations.</p>
* <span style="color: blue;">NO REGRETS BIAS</span> <p style="color:black;font-size:16px;text-align:left;"> The mechanism is optimised to reduce the risk of failing to activate when there is a shock, at the cost of having false alarms. `r mitigation_warranted `</p>

&nbsp;

_________

## About Metric Estimates

<p style="color:black;font-size:16px;text-align:left;"> The estimates in this report are shown as ranges called confidence intervals. The confidence intervals reflect the probabilistic nature of the estimates as well as the impact of data limitations (small or incomplete datasets, errors or imprecision in measurements, divergences between data sources, etc.) on the accuracy of estimates. They can inform decision-making by illustrating the most likely performance levels and by providing bounds for the possible albeit unlikely extremes.</p>

<br>

```{r explanatory_ci, echo=FALSE, out.width= "40%", out.extra='style="float:right; padding:10px"'}
explanatory_ci <- magick::image_read(explanatory_ci_plot_path)

explanatory_ci
```

* <p style="color:black;font-size:16px;text-align:left;"> The <span style="font-weight: bold;">central value</span> is the most likely performance; </p>
* <p style="color:black;font-size:16px;text-align:left;"> Most often the trigger will perform within the <span style="font-weight: bold;">darker inner range</span> (68% of the time);</p>
* <p style="color:black;font-size:16px;text-align:left;"> There is high confidence that the trigger will perform within the <span style="font-weight: bold;">full coloured (grey or blue) range</span> (95% of the time).</p>

<br>

## Colour Coding
<p style="color:black;font-size:16px;text-align:left;"> In the visualizations <span style="color: blue;font-weight: bold;">blue</span> represents activation likelihood, <span style="color: green;font-weight: bold;">green</span> represents desired outcomes (valid non-activations/activations) and <span style="color: red;font-weight: bold;"dou>red</span> represents undesired outcomes or errors.</p>

&nbsp;

## Trigger Mechanism Snapshot
<p style="color:black;font-size:16px;text-align:left;"> This table summarises the trigger mechanism and its performance in signaling severe shocks. It presents estimates of the likelihood that the threshold would be met in a given year as well as how often the trigger is expected to correctly recommend an activation or non-activation.</p>

&nbsp;
```{r mechanism-snapshot, include = TRUE}

snapshot_df <- data.frame(row_name = c(
                                       'Lead Time',
                                       'Data Source',
                                       'Historical Data Coverage',
                                       'Monitored by',
                                       'Probability of being met in a given year',
                                       'All Activations (%)',
                                       'All Shocks (%)'),
                       trigger_a = c(
                                     lead_time_a,
                                     data_source_a,
                                     hist_period_a,
                                     monitored_by_a,
                                     "<img src=plots/trimmed_readiness_atv.png>",
                                     "<img src=plots/trimmed_readiness_var_tradeoff_bar.png>",
                                     "<img src=plots/trimmed_readiness_det_tradeoff_bar.png>"),
                      trigger_b = c(
                                     lead_time_b,
                                     data_source_b,
                                     hist_period_b,
                                     monitored_by_b,
                                     "<img src=plots/trimmed_activation_atv.png>",
                                     "<img src=plots/trimmed_activation_var_tradeoff_bar.png>",
                                     "<img src=plots/trimmed_activation_det_tradeoff_bar.png>"),
                     trigger_c = c(
                                     lead_time_c,
                                     data_source_c,
                                     hist_period_c,
                                     monitored_by_c,
                                     "<img src=plots/trimmed_framework_atv.png>",
                                     "<img src=plots/trimmed_framework_var_tradeoff_bar.png>",
                                     "<img src=plots/trimmed_framework_det_tradeoff_bar.png>")
)

kbl(snapshot_df,
    format = 'html',
    escape = FALSE,
    row.names = FALSE, # omit row numbers
    col.names = c("","Readiness Trigger", "Activation Trigger", "Full Framework"),
    align = "rccc") %>%
  kable_minimal(c("hover", "striped")) %>%
  row_spec(c(1:5), extra_css = "line-height: 30px;") %>%
  row_spec(c(6:7), extra_css = "line-height: 85px;") %>%
  column_spec(column = 1, extra_css = "vertical-align: middle;") %>%
  column_spec(column = c(2:4), width = "28em") %>%
  pack_rows("Description", start_row = 1, end_row = 4, color = '#1bb580') %>%
  pack_rows("Programming and Funding", start_row = 5, end_row = 5, color = '#1bb580') %>%
  pack_rows("Performance", start_row = 6, end_row = 7
            , color = '#1bb580') %>%
  kableExtra::footnote(general_title = "Notes",
                       general = trigger_table_footnote)

```

&nbsp;

## Technical Team
<p style="color:black;font-size:16px;text-align:left;"> `r data_providers` have provided data and/or analytical support.

## Monitoring Process
<p style="color:black;font-size:16px;text-align:left;"> `r monitoring_process` </p>

&nbsp;
